Wrinkling à la Biot

### 2015-01-24 2014_10_23_memo.tex
Correct typo
Growth rate maximisation

\usepackage{siunitx}

@@

Mechanical equilibrium of a plate of thickness $h$, young modulus $E$ and Poisson ratio $\nu$ submitted to a compression load $\sigma_0$ along $x$ and a lateral load $q$ along $z$:
\begin{equation}
\frac{E h^3}{12(1-\nu^2)}\frac{\partial^4 w}{\partial w^4} + \sigma_0 h \frac{\partial^2 w}{\partial x^2} = q,
\end{equation}
where $w(x,t)$ is the deflection of the plate along $z$. It can be generalised as
\begin{equation}
B\frac{h^3}{12}\frac{\partial^4 w}{\partial w^4} + \sigma_0 h \frac{\partial^2 w}{\partial x^2} = -B_s \frac{\partial w}{\partial x},
\label{eq:mecheq}
\end{equation}
where $B=\frac{E}{1-\nu^2}$ and $B_s$ are the respective moduli (in \si{\pascal}) of the plate and the (effective) substrate. In general $B$ and $B_s$ are operators.

@@

In operator language
\begin{equation}
\frac{12\eta}{\tau} \frac{w}{H^3k^2} = -q
\end{equation}
which allows to identify $B_s = \frac{12\eta}{\tau H^3k^3}$. Since $B_s$ depends on $k$, we come back to Eq.~(\ref{eq:sigma0}) that we differentiate:
\begin{equation}
0 = \frac{\partial\sigma_0}{\partial k}
 = \frac{Bh^2 k}{6} - \frac{48}{hH^3}\frac{\eta}{\tau}\frac{1}{k^5},
\end{equation}
which yields 
\begin{equation}
k_P^6 = 2\times 12^2 \frac{1}{h^3H^3\Upsilon},
\end{equation}
with $\Upsilon = \frac{B\tau}{\eta}$ the viscoelastic factor. The dominant wavelength for Poiseuille model is thus
\begin{equation}
\lambda_P = \pi\sqrt{hH}\left(\frac{2}{9}\Upsilon\right)^{1/6}
\end{equation}

@@

Since we only consider sinusoidal variations $Q_1 + Q_2 = 0$. Using the expression of the Poiseuille flux Eq.~(\ref{eq:PoiseuilleFlux}), we obtain
\begin{equation}
H_1^3 \frac{\partial p_1}{\partial x} + H_2^3 \frac{\partial p_2}{\partial x} = 0,
\end{equation}
and thus
\begin{equation}
H_1^3 p_1 + H^2 p_2 = 0.
\label{eq:pressures}
\end{equation}

We now write the mass conservation in the lower viscous layer
\begin{equation}
\frac{\partial Q}{\partial x} + \frac{\partial w}{\partial t} + v = 0,
\label{eq:conservationDarcy}
\end{equation}
which is the same as Eq.~(\ref{eq:conservation}) with an added leak $v$ due to the porosity. $v$ can be expressed by the Darcy law
\begin{equation}
v = \frac{\alpha}{\eta} \frac{p_1-p_2}{h}.
\label{eq:Darcy}
\end{equation}

@@

The lateral load is
\begin{equation}
q = p_1-p_2 = \left(\frac{H_1}{H}\right)^3 p_1 = - \frac{\eta}{\tau} \frac{w}{\frac{H^3 k^2}{12} + \frac{\alpha}{h}},
\end{equation}
which allows to identify 
\begin{equation}
B_s = \frac{\eta}{\tau}\frac{1}{\frac{H^3 k^2}{12} + \frac{\alpha}{h}}.
\end{equation}

@@

>>>>

\usepackage{siunitx}
\usepackage{hyperref}

@@

Mechanical equilibrium of a plate of thickness $h$, young modulus $E$ and Poisson ratio $\nu$ submitted to a compression load $\sigma_0$ along $x$ and a lateral load $q$ along $z$:
\begin{equation}
\frac{E h^3}{12(1-\nu^2)}\frac{\partial^4 w}{\partial x^4} + \sigma_0 h \frac{\partial^2 w}{\partial x^2} = q,
\end{equation}
where $w(x,t)$ is the deflection of the plate along $z$. It can be generalised as
\begin{equation}
B\frac{h^3}{12}\frac{\partial^4 w}{\partial x^4} + \sigma_0 h \frac{\partial^2 w}{\partial x^2} = -B_s \frac{\partial w}{\partial x},
\label{eq:mecheq}
\end{equation}
where $B=\frac{E}{1-\nu^2}$ and $B_s$ are the respective moduli (in \si{\pascal}) of the plate and the (effective) substrate. In general $B$ and $B_s$ are operators.

@@

In operator language
\begin{equation}
\frac{12\eta}{\tau} \frac{w}{H^3k^2} = -q
\end{equation}
which allows to identify $B_s = \frac{12\eta}{\tau H^3k^3}$. Since $B_s$ depends on $k$, we come back to Eq.~(\ref{eq:sigma0}) that we differentiate:
\begin{equation}
0 = \frac{\partial\sigma_0}{\partial k}
 = \frac{Bh^2 k}{6} - \frac{48}{hH^3}\frac{\eta}{\tau}\frac{1}{k^5},
\end{equation}
which yields 
\begin{equation}
k_P^6 = 2\times 12^2 \frac{1}{h^3H^3\Upsilon},
\label{eq:Poiseuillek}
\end{equation}
with $\Upsilon = \frac{B\tau}{\eta}$ the viscoelastic factor. The dominant wavelength for Poiseuille model is thus
\begin{equation}
\lambda_P = \pi\sqrt{hH}\left(\frac{2}{9}\Upsilon\right)^{1/6}
\end{equation}

@@

Since we only consider sinusoidal variations $Q_1 + Q_2 = 0$. Using the expression of the Poiseuille flux Eq.~(\ref{eq:PoiseuilleFlux}), we obtain
\begin{equation}
H_1^3 \frac{\partial p_1}{\partial x} + H_2^3 \frac{\partial p_2}{\partial x} = 0,
\end{equation}
and thus
\begin{equation}
H_1^3 p_1 + H_2^3 p_2 = 0.
\label{eq:pressures}
\end{equation}

We now write the mass conservation in the lower viscous layer
\begin{equation}
\frac{\partial Q_1}{\partial x} + \frac{\partial w}{\partial t} + v = 0,
\label{eq:conservationDarcy}
\end{equation}
which is the same as Eq.~(\ref{eq:conservation}) with an added leak $v$ due to the porosity. $v$ can be expressed by the Darcy law
\begin{equation}
v = \frac{\alpha}{\eta} \frac{p_1-p_2}{h}.
\label{eq:Darcy}
\end{equation}

@@

The lateral load is
\begin{equation}
q = p_1-p_2 = \left(\frac{H_1}{H}\right)^3 p_1 = - \frac{\eta}{\tau} \frac{w}{\frac{H^3 k^2}{12} + \frac{\alpha}{h}},
\end{equation}
which allows to identify 
\begin{equation}
B_s = \frac{\eta}{\tau}\frac{1}{\frac{H^3 k^3}{12} + \frac{\alpha k}{h}}.
\end{equation}

@@

\section{Growth rate maximisation}

Rather than finding the mode that minimises the stress, it is possible to find the mode that has the maximum growth rate. Disclaimer: both methods always give the same result.

In the case of an elastic film on a viscous layer the mechanical equilibrium writes:
\begin{equation}
\left(\frac{B h^3}{12}k^4 - \sigma_0 h k^2\right) w = -\frac{12\eta}{H^3 k^2}\frac{\partial w}{\partial t},
\end{equation}
which is of the form $\tau \frac{\partial w}{\partial t}=w$, with
\begin{equation}
\tau = \frac{12^2\eta}{H^3 h (12 \sigma_0 k^4 - B h^2 k^6)}.
\end{equation}

To minimise $\tau$ we must have
\begin{equation}
0 = \frac{\partial}{\partial k}\left(12 \sigma_0 k^4 - B h^2 k^6\right) = 6k^3\left( 8 \sigma_0 - B h^2 k^2\right).
\end{equation}
Thus the minimum $\tau$ is
\begin{equation}
\tau_\text{min} = \frac{2\times12^2\eta}{H^3 h^3 k^6 B},
\end{equation}
from which we express the wave number
\begin{equation}
\frac{2\time 12}{k^6} = H^3 h^3 \frac{B\tau_min}{\eta} = H^3 h^3  \Upsilon,
\end{equation}
which is the same as Eq.~(\ref{eq:Poiseuillek}).
